Quantifying Stability and Change in Personal Culture Using Panel Data

### 2023-09-27 manuscript/manuscript.Rmd
color figs, inline, and endnotes.

---
title: "Quantifying Stability and Change in Personal Culture Using Panel Data"
author: 
- 
output:
  pdf_document:
    keep_tex: true
geometry: margin=1in 
mainfont: Times New Roman
fontsize: 12pt
header-includes: \usepackage{setspace}\doublespacing
bibliography: kileybib.bib
csl: asa.csl
abstract: "Recent work has produced conflicting interpretations on the question of whether adults undergo durable changes in their personal culture over time. This paper asserts that this divergence is less due to researchers' findings than the approach taken -- a tournament of models focusing on the presence or absence of change. To move beyond this, we reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference. In doing so, we compare how much cultural difference is explained by intrapersonal change and interpersonal differences. Our results harmonize recent findings by showing that, while most measures of personal culture show change over time, the variance explained by intrapersonal change is typically small compared to interpersonal differences at baseline. Quantifying these processes offers new ways to address theoretical questions of when and how their relative importance shifts. We examplify this approach by exploring how the importance of these processes differs between college graduates and the rest of the population."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
```

@@

```{=tex}
\begin{center}
<about here> 
Table 1: The Description of the Data Sources
\end{center}
```
## Life Course Adaption Model

@@

```{=tex}
\begin{center}
<about here>
Figure 1: Proportions of Systematic Variance in Personal Culture
\end{center}
```
Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change. Again, this is to be expected. Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture.

@@

```{=tex}
\begin{center}
<about here>
Figure 2: Regression Model Estimating $\omega$
\end{center}
```
Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$. We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change. The earlier a question was asked, even net of how long it was been asked, the lower the $\omega$.

@@

```{=tex}
\begin{center}
<about here>
Figure 3: Difference in $\omega$ Across College Graduates and Non-Graduates
\end{center}
```
There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., the difference is negative). While most of these differences are small in absolute terms, less than 2 percentage points, several are greater than 5 percentage points. Given the distribution observed in Figure 1 showing that the systematic variance attributable to intrapersonal change averages around 0.09, a 5 percentage point difference between groups is quite substantial.

@@

```{=tex}
\begin{center}
<about here>
Figure 4: Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture
\end{center}
```
On all eight questions presented in Figure 4, $\omega$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one. There are also large differences in $\omega$ values across questions for both groups. For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups. In other words, while people might vacillate on this question at random (37 percent of variance is residual for this question), there is functionally no evidence that people make systematic changes of opinion on this issue during the GSS panel.

@@

# Endnotes

@@

\newpage

# Tables and Figures

## Tables

kableExtra::kable(table, 
                  format = "latex", 
                  caption = "The Description of the Data Sources") |> 
  kable_styling(latex_options = c("scale_down", "HOLD_position"))

\newpage

## Figures

```{=latex}
\begin{figure}[ht]
\caption{Proportions of Systematic Variance in Personal Culture}
```
```{r fig1, fig.height=2}
knitr::include_graphics("../figures/figure_1_bw.png")
```

```{=latex}
\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively. See Supplemental Materials A for the full set of item values. 
\end{figure}
```
\newpage

```{=latex}
\begin{figure}[ht]
\caption{Regression Model Estimating $\omega$}
```
```{r fig2, out.width = "300px"}
knitr::include_graphics("../figures/figure_2.png")
```

```{=latex}
\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change. It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects. Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point. Survey indicators and item topics not shown. See Supplemental Materials B for the full set of coefficients.
\end{figure}
```
\newpage

```{=latex}
\begin{figure}[ht]
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
```
```{r fig3, out.width = "450px"}
knitr::include_graphics("../figures/figure_3_bw.png")
```

```{=latex}
\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates. Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change. The dashed red line marks 0 difference.
\end{figure}
```
\newpage

```{=latex}
\begin{figure}[ht]
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
```
```{r fig4, out.width = "450px"}
knitr::include_graphics("../figures/figure_4_bw.png")
```

```{=latex}
\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
\end{figure}
```

>>>>

---
title: "Quantifying Stability and Change in Personal Culture Using Panel Data"
author: 
- 
output:
  pdf_document:
    keep_tex: true
geometry: margin=1in 
mainfont: Times New Roman
fontsize: 12pt
header-includes:
  \usepackage{setspace}\doublespacing
  \usepackage{endnotes}
  \renewcommand{\enotesize}{\normalsize}
  \let\footnote=\endnote
bibliography: kileybib.bib
csl: asa.csl
abstract: "Recent work has produced conflicting interpretations on the question of whether adults undergo durable changes in their personal culture over time. This paper asserts that this divergence is less due to researchers' findings than the approach taken -- a tournament of models focusing on the presence or absence of change. To move beyond this, we reanalyze the data used previously in this debate with an approach that quantifies the processes of cultural difference. In doing so, we compare how much cultural difference is explained by intrapersonal change and interpersonal differences. Our results harmonize recent findings by showing that, while most measures of personal culture show change over time, the variance explained by intrapersonal change is typically small compared to interpersonal differences at baseline. Quantifying these processes offers new ways to address theoretical questions of when and how their relative importance shifts. We examplify this approach by exploring how the importance of these processes differs between college graduates and the rest of the population."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

@@

kableExtra::kable(table,
                  booktabs = TRUE,
                  format = "latex",
                  align = "llcc",
                  caption = "The Description of the Data Sources") |> 
  kable_styling(latex_options = c("scale_down", "HOLD_position"))

## Life Course Adaption Model

@@

```{=latex}
\begin{figure}[htp]
\begin{center}
\caption{Proportions of Systematic Variance in Personal Culture}
```
```{r fig1, out.width = "100%", fig.pos = "H"}
knitr::include_graphics("../figures/figure_1.png")
```
```{=latex}
\end{center}
\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences. See Supplemental Materials A for the full set of item values.}
\end{figure}
```

Across all questions, interpersonal differences explain a much larger share of the systematic variance in responses than intrapersonal change. Again, this is to be expected. Interpersonal differences capture not just pre-adult socialization, but all accumulated experiences up to the start of the panel that might have a bearing on personal culture.

@@

```{=latex}
\begin{figure}[htp]
\begin{center}
\caption{Regression Model Estimating $\omega$}
```
```{r fig2, out.width = "450px"}
knitr::include_graphics("../figures/figure_2.png")
```
```{=latex}
\end{center}
\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change. It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects. Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point. Survey indicators and item topics not shown. See Supplemental Materials B for the full set of coefficients.}
\end{figure}
```

Figure 2 shows that the more response options respondents are given and the larger the sample, the larger the values of $\omega$. We interpret these coefficients as suggesting greater resolution on a question makes it easier to detect and model change. The earlier a question was asked, even net of how long it was been asked, the lower the $\omega$.

@@

```{=latex}
\begin{figure}[hbt]
\begin{center}
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}
```
```{r fig3, out.width = "450px"}
knitr::include_graphics("../figures/figure_3.png")
```

```{=latex}
\end{center}
\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates. Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change. The dashed red line marks 0 difference.}
\end{figure}
```

There is a clear pattern in Figure 3: for more than 80 percent of GSS items explored here, intrapersonal change is a larger component of systematic variance for people without a college degree than for people with a college degree (i.e., the difference is negative). While most of these differences are small in absolute terms, less than 2 percentage points, several are greater than 5 percentage points. Given the distribution observed in Figure 1 showing that the systematic variance attributable to intrapersonal change averages around 0.09, a 5 percentage point difference between groups is quite substantial.

@@

```{=latex}
\begin{figure}[hbt]
\begin{center}
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}
```
```{r fig4, out.width = "450px"}
knitr::include_graphics("../figures/figure_4.png")
```

```{=latex}
\end{center}
\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
\end{figure}
```

On all eight questions presented in Figure 4, $\omega$ values are greater for people without a college degree than for people with a college degree, meaning intrapersonal change explains less systematic variance for people with a college degree than for people without one. There are also large differences in $\omega$ values across questions for both groups. For example, on the question of whether the government should try to solve more problems or leave those problems to be solved by private businesses ("government do more or less"), less than one percent of the systematic variance is attributable to intrapersonal change for both groups. In other words, while people might vacillate on this question at random (37 percent of variance is residual for this question), there is functionally no evidence that people make systematic changes of opinion on this issue during the GSS panel.

@@

\theendnotes

@@

### 2023-09-27 manuscript/manuscript.tex
color figs, inline, and endnotes.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}\doublespacing
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Quantifying Stability and Change in Personal Culture Using Panel Data},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Our method can also help address substantive questions about personal
change. To illustrate this, we investigate differences in the relative
importance of intrapersonal change on political items for people with
and without a college degree. We find that change is less pronounced for
college-educated respondents, suggesting that college solidifies
political dispositions rather than fostering an openness to new
information. This offers a novel empirical basis to theorize about the
role of college in shaping personal culture . In summary, we believe
that this approach has the potential to advance the theoretical debate
on the role of environments, social identities, and roles in
facilitating personal cultural change and stability.

Taken together, to goal of this paper is to advance---and, hopefully,
transcend---debates about \emph{whether} adults change. They do change,
at least a little, on most things. Our method can help quantify exactly
how much.

@@

Current debates often narrow down to ask whether people \emph{ever}
change, which is rarely theoretically interesting since the answer is
almost always ``yes.'' Over sufficient time and in a large enough
sample, researchers will likely observe some evidence for durable
changes in individuals. A lack of evidence for change may simply due to
a lack of clarity in the question asked or a low level of resolution in
the possible answers, or that the question simply was not asked for long
enough or to enough people. Conversely, finding evidence of change tells
us little about how much change is present in a population and how
meaningful that change is for social differences.

A more theoretically productive approach begins with a model assuming
that during an observed time period people might remain stable, might
change a little, or might undergo significant shifts in their personal
culture. Lersch's LCAM formally does this by modeling each individual as
following their own trajectory over time. But rather than debating
whether this model fits a particular data set better than a model that
assumes that people never change, researchers should quantify individual
changes over time and compare those to the differences that exist
between people at the panel's start. In other words, the focus should
shift from asking, ``do people change?'' to asking, ``what are the
relative contributions of change and initial differences toward
explaining differences in personal culture?''

In this approach, there are two metrics that each reflect a combination
of relevant theoretical processes. The first metric, interpersonal
differences at the start of the survey, reflects the accumulated
experiences of people prior to entering the panel survey that shape how
they answer questions, what Lersch calls ``early imprinting.'' While
commonly associated with experiences during an early formative period
that result in settled dispositions, this variation could also reflect
experiences that happen at any stage in life as long as they predate the
panel and consistently impact subsequent responses. For instance, for
those people entering the panel post-retirement, this ``imprinting''
might reflect this pivotal life transition. Consequently, this metric
also reflects variation in individuals' social roles or statuses at the
start of the panel that were important in shaping their dispositions.

The second metric, the amount of intrapersonal linear change people make
during the panel, captures durable changes in personal culture over
time, a set of processes collectively called ``persistent change'' or
``adaption'' by Lersch and ``active updating'' by Kiley and Vaisey.
Lersch attributes these changes to social triggers such as moving into a
new environment or adopting new social roles. Additionally, they might
reflect the diffusion of new cultural forms across social networks, cues
from political elites or otherwise culturally influential leaders, the
emergence of issues in politics or culture, or large-scale social
shifts.

A third metric, often called ``residual variance'' or ``fluctuation,''
accounts for the remainder of variance in peoples' responses over time.
This non-durable change emerges for a variety of reasons. For example,
people might not have clear personal cultural dispositions as measured
by surveys. Instead, they might internalize a broad set of
considerations and construct an opinion in the context of the survey
interview, with different considerations coming to the forefront of
their cognition each interview (Feldman and Zaller 1992; Tourangeau,
Rips, and Rasinski 2000; Zaller 1992). This variance in considerations
can produce variance in responses even if peoples' ``dispositions''
--the full set of considerations they could draw on-- remains stable
over time. Additionally, this variance can also include measurement
errors, such as misinterpreted questions, erroneous response selections,
or responses getting coded incorrectly. While this third process
reflects important aspects of personal culture, it does not directly
touch on the ongoing debate outlined here. Because of this we quantify
this component but focus principally on the other two.

@@

\hypertarget{analytical-strategy}{%
\section{Analytical Strategy}\label{analytical-strategy}}

@@

\begin{center}
<about here> 
Table 1: The Description of the Data Sources
\end{center}

@@

\begin{center}
<about here>
Figure 1: Proportions of Systematic Variance in Personal Culture
\end{center}

@@

\begin{center}
<about here>
Figure 2: Regression Model Estimating $\omega$
\end{center}

@@

\begin{center}
<about here>
Figure 3: Difference in $\omega$ Across College Graduates and Non-Graduates
\end{center}

@@

\begin{center}
<about here>
Figure 4: Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture
\end{center}

@@

\hypertarget{endnotes}{%
\section{Endnotes}\label{endnotes}}

@@

\newpage

\hypertarget{tables-and-figures}{%
\section{Tables and Figures}\label{tables-and-figures}}

\hypertarget{tables}{%
\subsection{Tables}\label{tables}}

\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l|l|l|r}
\hline
Country & Survey & Period & Outcomes\\
\hline
Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
\hline
Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
\hline
Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
\hline
Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
\hline
Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
\hline
United States & General Social Survey (GSS) & 2006-2012 & 183\\
\hline
United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
\hline
\end{tabular}}
\end{table}

\newpage

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}[ht]
\caption{Proportions of Systematic Variance in Personal Culture}

\begin{center}\includegraphics[width=20.83in]{../figures/figure_1_bw} \end{center}

\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences, respectively. See Supplemental Materials A for the full set of item values. 
\end{figure}

\newpage

\begin{figure}[ht]
\caption{Regression Model Estimating $\omega$}

\begin{center}\includegraphics[width=300px]{../figures/figure_2} \end{center}

\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change. It includes the log of the number of participants, the number of waves the variable was asked (ceiled at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects. Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point. Survey indicators and item topics not shown. See Supplemental Materials B for the full set of coefficients.
\end{figure}

\newpage

\begin{figure}[ht]
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}

\begin{center}\includegraphics[width=450px]{../figures/figure_3_bw} \end{center}

\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates. Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change. The dashed red line marks 0 difference.
\end{figure}

\newpage

\begin{figure}[ht]
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}

\begin{center}\includegraphics[width=450px]{../figures/figure_4_bw} \end{center}

\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Quantifying Stability and Change in Personal Culture Using Panel Data},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

In addition to advancing a theoretical debate, our method can also help
address substantive questions. To illustrate how it might be used by a
researcher with a specific topical interest, we investigate differences
in the relative importance of intrapersonal change for people with and
without a college degree. We find a smaller amount of systematic change
among college-educated respondents, suggesting that college solidifies
one's personal culture rather than fostering an openness to new
information. This offers a novel empirical basis to theorize about the
role of university education in shaping personal culture.

Taken together, to goal of this paper is to advance---and, hopefully,
transcend---debates about \emph{whether} adults change. People do
change, at least a little, on most things. Our method can help quantify
exactly how much.

@@

The current debate can be simplified to asking whether people
\emph{ever} change. This is rarely theoretically interesting since the
answer is almost always ``yes.'' Over sufficient time and in a large
enough sample, researchers will likely observe some evidence for durable
individual change. A lack of evidence for change may be due to a poor
survey question, a low resolution in response options, or that the
question simply was not asked for long enough or to enough people.
Conversely, finding evidence for some change tells us only that it was
``not zero.'' From that alone, we learn little about how much change has
happened in a population.

A more theoretically productive approach begins with a model assuming
that during an observed time period people might remain stable, might
change a little, or might undergo significant shifts in their personal
culture. Lersch's LCAM does this by modeling each individual as
following their own linear trajectory. But rather than debating whether
this model fits a particular data set better than a model that assumes
that people never change, researchers should quantify these individual
changes and compare those to the differences that exist between people
at the panel's start. In other words, the focus should shift from
asking, ``do people change?'' to asking, ``what are the relative
contributions of change and initial differences for explaining
differences in personal culture?''

In this approach, there are two metrics that each reflect a combination
of relevant theoretical processes. The first metric, interpersonal
differences at the start of the survey, reflects the accumulated
experiences of people prior to entering the panel survey that shape how
they answer questions. Lersch calls this ``early imprinting.'' While
commonly associated with experiences during a formative period that
result in settled dispositions, this variation could also reflect
experiences that happen at any stage in life as long as they predate the
panel and consistently affect subsequent responses. For instance, for
those people entering the panel post-retirement, this ``imprinting''
might reflect this pivotal life transition. Consequently, this metric
also reflects variation in individuals' social roles or statuses at the
start of the panel that were important in shaping their dispositions.

The second metric, the amount of intrapersonal linear change people make
during the panel, captures durable changes in personal culture over
time, a set of processes collectively called ``persistent change'' or
``adaption'' by Lersch and ``active updating'' by Kiley and Vaisey.
Lersch attributes these changes to social triggers such as moving into a
new environment or adopting new social roles (although he never measures
them directly). Additionally, they might reflect the diffusion of new
cultural forms across social networks, cues from political elites or
otherwise culturally influential leaders, the emergence of issues in
politics or culture, or large-scale social shifts.

A third metric, often called ``residual variance'' or ``fluctuation,''
accounts for the remainder of variance in peoples' responses. This
non-durable change emerges for a variety of reasons. For example, people
might not have clear a clear disposition on a particular item as it is
asked. Instead, they might internalize a broad set of considerations and
construct an opinion in the context of the survey interview, with
different considerations coming to the forefront of their cognition each
interview (Feldman and Zaller 1992; Tourangeau, Rips, and Rasinski 2000;
Zaller 1992). Additionally, this variance can also include measurement
error, such as misinterpreted questions, erroneous response selections,
or responses getting coded incorrectly. While this third metric at least
partially reflects important processes of personal culture, it does not
directly touch on the ongoing debate outlined here. Because of this we
quantify this component but focus principally on the other two.

@@

\hypertarget{analytic-strategy}{%
\section{Analytic Strategy}\label{analytic-strategy}}

@@

\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llcc}
\toprule
Country & Survey & Period & Outcomes\\
\midrule
Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
\addlinespace
United States & General Social Survey (GSS) & 2006-2012 & 183\\
United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
\bottomrule
\end{tabular}}
\end{table}

@@

\begin{figure}[htp]
\begin{center}
\caption{Proportions of Systematic Variance in Personal Culture}

\includegraphics[width=1\linewidth]{../figures/figure_1}

\end{center}
\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences. See Supplemental Materials A for the full set of item values.}
\end{figure}

@@

\begin{figure}[htp]
\begin{center}
\caption{Regression Model Estimating $\omega$}

\includegraphics[width=450px]{../figures/figure_2}

\end{center}
\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change. It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects. Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point. Survey indicators and item topics not shown. See Supplemental Materials B for the full set of coefficients.}
\end{figure}

@@

\begin{figure}[hbt]
\begin{center}
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}

\includegraphics[width=450px]{../figures/figure_3}

\end{center}
\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates. Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change. The dashed red line marks 0 difference.}
\end{figure}

@@

\begin{figure}[hbt]
\begin{center}
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}

\includegraphics[width=450px]{../figures/figure_4}

\end{center}
\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
\end{figure}

@@

\theendnotes

@@
